---
title: "The Standard Normal Distribution (N(0,1)) and Z-scores"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
```

## Site I

```{r}
mu_1 <- 200
sigma_1 <- 6
mu_2 <- 210
sigma_2 <- 1
```

95% of the Martians at Site I ($\mu$ = `r mu_1` $\sigma$ = `r sigma_1`) are expected to have a height between `r mu_1 + c(-1.96,1.96)*(sigma_1)` 

## Site II

95% of the Martians at Site II ($\mu$ = `r mu_2` $\sigma$ = `r sigma_2`) are expected to have a height between `r mu_2 + c(-1.96,1.96)*(sigma_2)` 

## Diagram

"one value, two curves" - One country, two policies

```{r}
# Number of points on the distribution to calculate:
n_points <- 5000

# Construct plot
base_plot <- ggplot() + 
  geom_function(aes(color = "Site I"), fun = dnorm, args = list(mean = mu_1, sd = sigma_1), n = n_points) +
  geom_function(aes(color = "Site II"), fun = dnorm, args = list(mean = mu_2, sd = sigma_2), n = n_points) +
  coord_cartesian(expand = 0) +
  xlim(180, 220) +
  labs(color = "Location", x = "Height (cm)") +
  theme_classic() +
  theme(line = element_blank())

# Display plot:
#base_plot
```

Plotting the Mystery Martian:

```{r}
base_plot +
  geom_vline(aes(xintercept = 215))
```

## Calculating Z-scores

### Site I 

Calculating Z--scores for the Martian height at site I.

| $x_i (cm)$  | `r mu_1 - 3*sigma_1` | `r mu_1 - 2*sigma_1` | `r mu_1 - sigma_1` | `r mu_1` | `r mu_1 + sigma_1` | `r mu_1 + 2*sigma_1` | `r mu_1 + 3*sigma_1` |
|-------------|-----------------------------|-----------------------------|---------------------------|----------------|---------------------------|-----------------------------|-----------------------------| 
| $x_i - \mu$ |         -18                    |        -12                     |           -6                |         0       |              6             |               12              |                  18           |  
| $z_i$       |           -3                  |           -2                  |           -1                |         0       |             1              |              2               |                3             |

The Z--score for the Mystery martian at site I is: (215 - `r mu_1`)/`r sigma_1` = `r  (215 - mu_1)/sigma_1`.

### Site II

Calculating Z--scores for the Martian height at Site II

| $x_i (cm)$  | `r mu_2 - 3*1` | `r mu_2 - 2*1` | `r mu_2 - 1` | `r mu_2` | `r mu_2 + 1` | `r mu_2 + 2*1` | `r mu_2 + 3*1` |
|-------------|----------------------|----------------------|--------------------|----------------|--------------------|----------------------|----------------------| 
| $x_i - \mu$ |         -3             |            -2          |             -1      |      0          |                1    |           2           |           3           |  
| $z_i$       |           -3                  |           -2                  |           -1                |         0       |             1              |              2               |                3             |

The Z--score for the Mystery martian at site II is: (215 - `r mu_2`)/`r sigma_2` = `r  (215 - mu_2)/sigma_2`.

Z--scores are a _standardized_ score that tells us how many standard deviations (i.e. $\sigma$) any given value $x_i$ is away from the mean (i.e. $\mu$).

## The Standard Normal Distribution

Z--Scores are described by the Standard Normal Distribution, N($\mu$ = 0, $\sigma$ = 1). Thus an _extreme_ value is outside the interval [-1.96, 1.96] since values outside this range are less than 5% likely to be observed?

Z--Scores are our first experience with one of the most important equations in all of statistics: **signal:noise**. Here, signal: $x_i - \mu$, noise: $\sigma$.

"one curve, two values" - Lord of the rings model "One curve to rule them all".

```{r}
# Number of points on the distribution to calculate:
n_points <- 5000

# Construct plot
base_plot <- ggplot() + 
  geom_function(aes(color = "The Standard\nNormal\nDistribution"), fun = dnorm, args = list(mean = 0, sd = 1), n = n_points) +
  geom_vline(aes(xintercept = 2.5, color = "Z--score for\nMystery Martian\nat Site I")) +
  geom_vline(aes(xintercept = 5.0, color = "Z--score for\nMystery Martian\nat Site II")) +

  coord_cartesian(expand = 0) +
  xlim(-6, 6) +
  labs(color = "", x = "Z-score") +
  theme_classic() +
  theme(line = element_blank())

# Display plot:
base_plot
```



