---
title: "3 - Using functions"
author: "A bilingual case study"
---

```{r}
# If necessary:

# Load reticulate
# library(reticulate)

# Activate a virtual environment
# use_miniconda("r-reticulate", required = TRUE) 
```

Check active build:

```{r}
py_config()
```

We have previously installed `scikit-learn` in the `r-reticulate` conda env. Let's look at a hello world example.

First, import python package modules:

```{python}
from sklearn import datasets
from sklearn.svm import SVC
```

load the data:

```{python}
iris = datasets.load_iris()
iris.data[:6]
```

## Hello, SVC

Create an instance of the SVC, _Support Vector Classification_, class.

```{python}
clf = SVC()
```

Call the fit and predict methods on the target, using target names

```{python}
clf.fit(iris.data, iris.target_names[iris.target])
clf.predict(iris.data[:3])
```

## Accessing python functions

Note: This is completely meaningless since were predicting the exact same data we used to train the model. It's just an example of how to get things working, not a complete ML case study.

```{r}

# Predict classes
py$clf$predict(r_to_py(iris[-5]))
```
