---
title: 'Module 1: Introduction to Machine Learning'
output:
  html_document:
    df_print: paged
---

## Prerequisites

```{r}
# common data wrangling and visualization
# install.packages("tidyverse")
# install.packages("vip")

# modeling
# install.packages("tidymodels")
```

```{r}
packageVersion("tidymodels")

library(tidymodels)
```

## Boston housing

```{r}
# access data
boston <- readr::read_csv("./00-data/ML/boston.csv") 

# initial dimension
dim(boston)

# features
dplyr::select(boston, -cmedv)

# response variable
head(boston$cmedv)
```

## Pima Indians Diabetes

```{r}
# access data
pima <- readr::read_csv("./00-data/ML/pima.csv") 

# initial dimension
dim(pima)

# features
dplyr::select(pima, -diabetes)

# response variable
head(pima$diabetes)
```

## Iris flowers

```{r}
# access data
iris <- readr::read_csv("./00-data/ML/iris.csv") 

# initial dimension
dim(iris)

# features
dplyr::select(iris, -Species)

# response variable
head(iris$Species)
```

## Ames housing

```{r}
# access data
ames <- readr::read_csv("./00-data/ML/ames.csv") 

# initial dimension
dim(ames)

# features
dplyr::select(ames, -Sale_Price)

# response variable
head(ames$Sale_Price)
```

## Attrition

```{r import-attrition-data}
# access data
attrition <- readr::read_csv("./00-data/ML/attrition.csv") 

# initial dimension
dim(attrition)

# features
dplyr::select(attrition, -Attrition)

# response variable
head(attrition$Attrition)
```

## Exercises

1. Identify four real-life applications of supervised and unsupervised problems.
   - Explain what makes these problems supervised versus unsupervised.
   - For each problem identify the target variable (if applicable) and potential
     features.

2. Identify and contrast a regression problem with a classification problem.
   - What is the target variable in each problem and why would being able to
     accurately predict this target be beneficial to society?
   - What are potential features and where could you collect this information?
   - What is determining if the problem is a regression or a classification
     problem?

3. Identify three open source data sets suitable for machine learning (e.g.,
   https://bit.ly/35wKu5c).
   - Explain the type of machine learning models that could be constructed from
     the data (e.g., supervised versus unsupervised and regression versus
     classification).
   - What are the dimensions of the data?
   - Is there a code book that explains who collected the data, why it was
     originally collected, and what each variable represents?
   - If the data set is suitable for supervised learning, which variable(s) could
     be considered as a useful target? Which variable(s) could be considered as
     features?

4. Identify examples of misuse of machine learning in society. What was the
   ethical concern?
