---
title: "Regularized Regression"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Prerequisites

```{r}
# Helper packages
library(tidyverse) # general data munging & visualization

# Modeling packages
library(tidymodels)

# Model interpretability packages
library(vip)      # for variable importance
```

```{r 06-ames-train}
# Stratified sampling with the rsample package
ames <- AmesHousing::make_ames()
set.seed(123)  # for reproducibility
split  <- initial_split(ames, prop = 0.7, strata = "Sale_Price")
ames_train  <- training(split)
ames_test   <- testing(split)
```


# Implementation

## Ridge

```{r}
# Step 1: create ridge model object
ridge_mod <- linear_reg(penalty = 1, mixture = 0) %>%
  set_engine("glmnet")

# Step 2: create model & preprocessing recipe
model_recipe <- recipe(
    Sale_Price ~ Gr_Liv_Area + Year_Built + Garage_Cars + Garage_Area, 
    data = ames_train
  ) %>%
  step_normalize(all_predictors())
  
# Step 3: fit model workflow
ridge_fit <- workflow() %>%
  add_recipe(model_recipe) %>%
  add_model(ridge_mod) %>%
  fit(data = ames_train)

# Step 4: extract and tidy results
ridge_fit %>%
  pull_workflow_fit() %>%
  tidy()

# 1 (Intercept)  180923.       1
# 2 Gr_Liv_Area   37441.       1
# 3 Year_Built    22587.       1
# 4 Garage_Cars   10825.       1
# 5 Garage_Area   11495.       1

```

## Lasso

```{r}
# Step 1: create ridge model object
lasso_mod <- linear_reg(penalty = 1, mixture = 1) %>%
  set_engine("glmnet")
  
# Step 3: fit model workflow
lasso_fit <- workflow() %>%
  add_recipe(model_recipe) %>%
  add_model(lasso_mod) %>%
  fit(data = ames_train)

# Step 4: extract and tidy results
lasso_fit %>%
  pull_workflow_fit() %>%
  tidy()

# 1 (Intercept)  180923.       1
# 2 Gr_Liv_Area   40301.       1
# 3 Year_Built    24131.       1
# 4 Garage_Cars    9183.       1
# 5 Garage_Area   11269.       1
```

## Elastic net

```{r}
# Step 1: create ridge model object
en_mod <- linear_reg(penalty = 1, mixture = 0.5) %>%
  set_engine("glmnet")
  
# Step 3: fit model workflow
en_fit <- workflow() %>%
  add_recipe(model_recipe) %>%
  add_model(en_mod) %>%
  fit(data = ames_train)

# Step 4: extract and tidy results
en_fit %>%
  pull_workflow_fit() %>%
  tidy()

# 1 (Intercept)  180923.       1
# 2 Gr_Liv_Area   40200.       1
# 3 Year_Built    24094.       1
# 4 Garage_Cars    9267.       1
# 5 Garage_Area   11309.       1
```


# Tuning

## Tuning regularization strength

```{r}
# create linear model object
ridge_mod <- linear_reg(penalty = tune(), mixture = 0) %>%
  set_engine("glmnet")

# create k-fold cross validation object
folds <- vfold_cv(ames_train, v = 5)

# create our model recipe with a tuning option for number of components
model_recipe <- recipe(Sale_Price ~ ., data = ames_train) %>%
  step_normalize(all_numeric_predictors()) %>%
  step_dummy(all_nominal_predictors())

# create a hyper parameter tuning grid for penalty
hyper_grid <- grid_regular(penalty(range = c(-10, 5)), levels = 50)

# train our model across the hyper parameter grid
set.seed(123)
results <- tune_grid(ridge_mod, model_recipe, resamples = folds, grid = hyper_grid)

# get best results
show_best(results, metric = "rmse")

# 30850

```


## Tuning regularization type & strength

```{r}
# create linear model object
mod <- linear_reg(penalty = tune(), mixture = tune()) %>%
  set_engine("glmnet")

# create k-fold cross validation object
folds <- vfold_cv(ames_train, v = 5)

# create our model recipe with a tuning option for number of components
model_recipe <- recipe(Sale_Price ~ ., data = ames_train) %>%
  step_normalize(all_numeric_predictors()) %>%
  step_dummy(all_nominal_predictors())

# create a hyper parameter tuning grid for penalty
hyper_grid <- grid_regular(
  penalty(range = c(-10, 5)), 
  mixture(), 
  levels = c(mixture = 5, penalty = 50)
  )

# train our model across the hyper parameter grid
set.seed(123)
results <- tune_grid(mod, model_recipe, resamples = folds, grid = hyper_grid)

# get best results
show_best(results)
```


# Feature importance

```{r}
# identify best model
lowest_rmse <- results %>%
  select_best("rmse")

# extract best model workflow
final_model <- finalize_workflow(
  workflow() %>% add_recipe(model_recipe) %>% add_model(ridge_mod), 
  lowest_rmse)

# extract feature importance for top 25 most influential features
top_25_features <- final_model %>%
  fit(ames_train) %>%
  pull_workflow_fit() %>% 
  vi(lambda = lowest_rmse$penalty) %>%
  mutate(
    Importance = abs(Importance),
    Variable = fct_reorder(Variable, Importance)
  ) %>%
  top_n(25, wt = Importance)
```

```{r}
ggplot(top_25_features, aes(x = Importance, y = Variable, fill = Sign)) +
  geom_col() +
  scale_x_continuous(expand = c(0, 0)) +
  labs(y = NULL)
```


# Exercises

Using the boston housing dataset:

1. Apply a ridge model with glmnet with `medv` being the response variable.
   - What is the minimum MSE?
   - What is the minimum MSE within 1 standard error?
   - What are the lambda values for these MSEs?
2. Apply a lasso model with glmnet.
   - What is the minimum MSE?
   - What is the minimum MSE within 1 standard error?
   - What are the lambda values for these MSEs?
3. Perform a grid search across alpha parameter values ranging between 0–1.
   - What is the optimal alpha and lambda values?
   - What is the MSE and RMSE for this optimal model?
   - How does it compare to your previous models?
4. Plot the top 10 most influential features. Do these features have positive or negative impacts on your response variable?
