---
title: "Interactive Markdown Exercise 1"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
```

# An interactive ggplot

Add the code from our tutorial to make an interactive document:

Additional exercise: Also change the colors in some way

SSee help page for `shiny::selectInput()`

Input elements include (click for online R helpdocum:

[`selectInput()`](https://www.rdocumentation.org/packages/shiny.semantic/versions/0.4.3/topics/selectInput)

actionButton(), checkboxGroupInput(), checkboxInput(), dateInput(), dateRangeInput(), fileInput(), numericInput(), passwordInput(), radioButtons(), sliderInput(), submitButton(), textAreaInput(), textInput(), [varSelectInput()](http://127.0.0.1:50605/help/library/shiny/help/varSelectInput)


```{r echo = FALSE}

inputPanel({
  radioButtons("shape", label = "Circle:",
               c("Filled" = 16,
                 "Hollow" = 1))
})


inputPanel({
  sliderInput("col", label = "Choose color number", 
              min = 0,
              max = 100, 
              value = 1)
})


renderPlot({
  ggplot(mtcars, aes(mpg, wt)) +
    geom_point(shape = as.numeric(input$shape), 
               color = input$col,
               alpha = 0.75, size = 10)
})

```


