---
title: "Data Science Immersive: Data Visualization"
subtitle: "Introduction to ggplot2"
author: "Rick Scavetta, for Misk Skills"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    toc_depth: 3
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is the practical component of the **Data Visualization** module of
the Misk Skills Data Science Immersive

To use this markdown document, make sure you have installed `ggplot2`
(`munsell`, `RColorBrewer` and `Hmisc` should be installed as
dependencies). You can click on the knit button to execute all the code
and produce an HTML output, which is already provided for you in the
files pane to the right.

```{r}
# load packages, (pre-installed with tidyverse)
library(ggplot2)

# These are not strictly necessary, but we'll use them for looking at some specific examples 
library(munsell)
library(RColorBrewer)
library(Hmisc)
```

We'll begin by looking at accessing colors and then move onto `ggplot2`.

# Using Color

There many packages that provide colour palettes in R. The most widely
used and fundamental is `RColorBrewer`. These pallets come from Cynthia
Brewer, a cartographer who developed the pallets for her map-making
colleagues. They are a good starting point for us and well integrated
into ggplot2 and you can find a handy web-site summary
[here](https://colorbrewer2.org).

The *sequential* color palettes:

```{r}
display.brewer.all(type = "seq")
```

The *qualitative* color palettes:

```{r}
display.brewer.all(type = "qual")
```

The *divergent* color palettes:

```{r}
display.brewer.all(type = "div")
```

To view all the available pallets use:

```{r eval = FALSE}
display.brewer.all()
```

To access a given number of colors in a specific pallet use:

```{r}
display.brewer.pal(9, "Blues")
```

If you use

```{r}
brewer.pal(9, "Blues")
```

you'll get a character vector of the hex codes (discussed in the ebook).
Just like any other vector in R, we can index this using `[]`. Here,
we'll just take three shades of blue and assign them to the object
`myBlues`.

```{r}
myBlues <- brewer.pal(9, "Blues")[c(4,6,8)]
```

Once you have these hex codes, You can use them in all and graphics to
maintain consistency in your color palette across your presentation or
paper. If you want to view what colour the hex code is, the munsell
package has a really nice function:

```{r}
plot_hex(myBlues)
```

# ggplot2: The first three essential layers

Now that we have an idea of how to obtain color values, let's dig into
`ggplot2`. In the case studies I presented `ggplot2` as being comprised
of seven layers.

## Layer 1: The data layer

I discussed the data layer in the first case study looking at peppermint
sugar content. We need each variable on our plot to be its own variable
in our data site.

There are many built in datasets in R. A really good starting point for
understanding `ggplot2` is the `mtcars` data set. It describes the
properties of 32 engines from cars in 1973. The *mt* in `mtcars` stands
for *Motor Trends*, a magazine that apparently was ahead of the game in
providing detailed engine metrics.

The dataset contains the following variables:

| Variable | Description                              |
|----------|------------------------------------------|
| mpg      | Miles/(US) gallon                        |
| cyl      | Number of cylinders                      |
| disp     | Displacement (cu.in.)                    |
| hp       | Gross horsepower                         |
| drat     | Rear axle ratio                          |
| wt       | Weight (1000 lbs)                        |
| qsec     | 1/4 mile time                            |
| vs       | Engine (0 = V-shaped, 1 = straight)      |
| am       | Transmission (0 = automatic, 1 = manual) |
| gear     | Number of forward gears                  |
| carb     | Number of carburetors                    |

You can access the dataset directly from base package.

```{r}
mtcars
```

## Layers 2: Aesthetics

The second essential layer is the aesthetics layer. In the course, we
referred to these as the *encoding* elements. Other names for aesthetics
are *scales* or *axes*. e.g. we typically refer to the X axis, but there
are many other aesthetics which work in the same way. Let's begin with
the X and Y axes.

```{r}
# Univariate plot - Histogram
ggplot(mtcars, aes(x = mpg))
```

```{r}
# Bi-/Multi-variate plot - Scatter plot
ggplot(mtcars, aes(x = wt, y = mpg))
```

Notice that in the above plots, we have *literally* what we ask for: a
histogram with an X axis and a scatter plot with X and Y axes. There are
no visuals, i.e. geometries, since they need to be added as a separate
layer. We'll get to that soon.

The first arguments of `aes()` are `x` and `y`, so we'll typically leave
them out. Other common aesthetics are:

-   `color`

    -   Also, `colour`

-   `fill`

-   `shape`

-   `size`

-   `linetype`

You'll often see `color` called as as `col`, using partial name
matching, to call the argument. I tend to do this, but it's a bad habit.
The best practice is to call the entire name.

## Layers 3: Geometries

The third essential layer is the geometries layer. Here we specify what
we actually want to see. You'll see that the naming convention of the
functions in `ggplot2` are very well organized! All the geometries have
the form `geom_*()`. You can see a list by typing `geom_` in an R script
document and pressing tab for auto complete. There are currently 48
geometries:

| geom\_\* |            |           |           |            |          |        |
|----------|------------|-----------|-----------|------------|----------|--------|
| abline   | contour    | dotplot   | jitter    | pointrange | ribbon   | spoke  |
| area     | count      | errorbar  | label     | polygon    | rug      | step   |
| bar      | crossbar   | errorbarh | line      | qq         | segment  | text   |
| bin2d    | curve      | freqpoly  | linerange | qq_line    | sf       | tile   |
| blank    | density    | hex       | map       | quantile   | sf_label | violin |
| boxplot  | density2d  | histogram | path      | raster     | sf_text  | vline  |
| col      | density_2d | hline     | point     | rect       | smooth   |        |

For example, to make a histogram, we'll use:

```{r}
# Univariate plot - Histogram
ggplot(mtcars, aes(mpg)) +
  geom_histogram()
```

But for scatter plots, there is no scatter plot geom. We just need to
have continuous X and Y axes and apply a point geom layer:

```{r}
# Bi-/Multi-variate plot - Scatter plot
ggplot(mtcars, aes(wt, mpg)) +
  geom_point()
```

These are very basic parts, so let's spruce them up. Noticed that our
histogram automatically defaults to giving us 32 equally sized bins on
the X axis. `geom_histogram()` bins a continuous variable, counts the
number of observations in each bin, and then maps that variable onto a Y
aesthetic. It's likely that we'll want to have more intuitive bin sizes,
in this case 1 unit. It's also typical that the tick marks on the X axis
fall *between* the bars, and *not* in the middle. This tells us that
each bar is a range of values and not a discrete value.

```{r}
# Univariate plot - Histogram
ggplot(mtcars, aes(mpg)) +
  geom_histogram(binwidth = 1, boundary = 1)
```

For the scatter plots, we may want to adjust the geom's *attributes*:

```{r}
# Bi-/Multi-variate plot - Scatter plot
ggplot(mtcars, aes(wt, mpg)) +
  geom_point(
    shape = 16, 
    alpha = 0.6,
    size = 3,
    # fill = "pink"
    color = "green"
  )
```

The first argument, `shape` specifies the *shape* to use, as a built-in
R integer code. I find that the most common & useful values for the
`shape` argument are:

+---+------------------------------------+-----------------------------+
| V | Output                             | Aesthetic/Attribute         |
| a |                                    | argument                    |
| l |                                    |                             |
| u |                                    |                             |
| e |                                    |                             |
+===+====================================+=============================+
| ` | A single pixel, the smallest       | `color`                     |
| " | possible point.                    |                             |
| . |                                    |                             |
| " |                                    |                             |
| ` |                                    |                             |
+---+------------------------------------+-----------------------------+
| ` | Hollow circles                     | `color`                     |
| 1 |                                    |                             |
| ` |                                    |                             |
+---+------------------------------------+-----------------------------+
| ` | Filled circles, no outline (my     | `color`                     |
| 1 | favorite)                          |                             |
| 6 |                                    |                             |
| ` |                                    |                             |
+---+------------------------------------+-----------------------------+
| ` | Filled circle, with an outline of  | `color`                     |
| 1 | the *same* colour                  |                             |
| 9 |                                    |                             |
| ` |                                    |                             |
+---+------------------------------------+-----------------------------+
| ` | Filled circle, 2/3 the size of     | `color`                     |
| 2 | `19`                               |                             |
| 0 |                                    |                             |
| ` |                                    |                             |
+---+------------------------------------+-----------------------------+
| ` | Filled circle, with an outline of  | `color` the outline, `fill` |
| 2 | a *different* colour               | for the inside.             |
| 1 |                                    |                             |
| ` |                                    |                             |
+---+------------------------------------+-----------------------------+

The `fill` aesthetic and attribute typically refers to the *inside* of a
geometry, as we'll see later. However, An exception to the rule is with
points. Here, we typically use the `color` aesthetic, except in the case
where a circle has an outline which we see in shape `21`.

Recall that *alpha* refers to transparency. For `alpha`, `1` = 100%
opaque and `0` = 0% opaque (i.e. 100% transparent). I typically use a
value between 0.3 - 0.7.

The `size` argument adjusts the points size.

Returning to our discussion of over-plotting in the video course, in
this case I would use the above corrections (solid, slightly transparent
points) *or* I would go with hollow, opaque points.

```{r}
# Bi-/Multi-variate plot - Scatter plot
ggplot(mtcars, aes(wt, mpg)) +
  geom_point(shape = 1)
```

These argument names are the same names as the *aesthetics* I mentioned
earlier! This is common point of confusion, so note the distinction
between:

-   **Aesthetics**, which are arguments in the aesthetics layer, defined
    inside `aes()`, and are used to *map* a variable onto a scale, and
-   **Attributes**, which are arguments in the geom later, and are used
    to *set* a value.

The terms *scale*, *aesthetic*, and *axis* are commonly used
interchangeably, but at specific points in making the plot, so don't let
this confuse you.

Let's add another aesthetic to see what I mean:

```{r}
# Bi-/Multi-variate plot - Scatter plot
ggplot(mtcars, aes(wt, mpg, color = cyl)) +
  geom_point(shape = 16, alpha = 0.6,
             size = 3)
```

The points are now colored according to `cyl`, i.e. we *mapped* `cyl`
onto `color`. Unfortunately, it's the wrong colour scale. It's a
sequential, continuous color scale, but should be a qualitative color
scale. You may recall from the data analysis class that categorical
variables are defined as factors in R. We can set this on the fly in our
plot.

```{r}
# Bi-/Multi-variate plot - Scatter plot
ggplot(mtcars, aes(wt, mpg, col = factor(cyl))) +
  geom_point(shape = 16, alpha = 0.6,
             size = 3)
```

That's better. We'll change it permanently in the dataset later on.

So what happens if we set a colour aesthetic *and* a colour attribute?

```{r}
# Bi-/Multi-variate plot - Scatter plot
ggplot(mtcars, aes(wt, mpg, col = factor(cyl))) +
  geom_point(shape = 16, alpha = 0.6,
             size = 3, color = "pink")
```

Attributes beat out aesthetics! This is another really common problem,
but if we are aware of this behavior we can use it to good effect.

## Positions

So far our plots have used position "identity", Which means don't do
anything, just put the geometry exactly where the data says to put it.
That's the default for `geom_point()`, which kind of makes sense. But
that doesn't make sense for things like bar plots.

Let's take a look. Our goal is to make a plot of this contigency table.
It tells us the number of counts for each transmission (`am`) type, 0 &
1, in each cylinder (`cyl`) class, 4, 6, & 8.

```{r}
table(mtcars$am, mtcars$cyl)
```

Remember, `ggplot2` requires a dataframe as input, so we may be tempted
to convert the above table to a dataframe. We have the raw values, but
what we want to plot is the absolute count of the number of
observations. We could calculate the number of observations in each
group ourselves, but we can just let ggplot2 do all of that for us.

We're going to make a couple versions of this plot, so I'm going to
assign the data and aesthetics layers to an object, called `g`. You'll
see this a lot in ggplot2 programming. Sometimes the object names are
one letter long, which is typically pretty bad style but somehow people
get away with it in this case, mostly for the sake of convenience.
Typical letters are `g`, `p`, & `q`. Although I'm introducing this
concept here, since it makes this example convenient, I want to
emphasize that you **do not** need to do this when just making a single
plot. It has lots of very useful applications for when your code gets
more complicated. You'll see it in help pages and tutorials, so it's
good to be aware of this practice.

```{r}
# base layers
g <- ggplot(mtcars, aes(factor(cyl), fill = factor(am)))

g

```

`geom_bar()` will apply a count statistic to our data, which is
basically what `table()` did.

```{r}
# Default bar plot
g +
  geom_bar()
```

The default position is `"stack"`, which can be specified as:

```{r}
# Which is...
g +
  geom_bar(position = "stack")

```

```{r}
g +
  geom_bar(position = "fill") +
  scale_y_continuous("Proportion") +
  coord_flip()
```

Position `"stack"` is pretty problematic because there is inherent
confusion if the bars are actually *stacked* or if they are
*overlapping* each other. It sounds strange, but actually, it's not
obvious to the reader which one it is, so we want to avoid this
ambiguity. Position `"dodge"` takes care of this for us:

```{r}

# position "dodge"
g +
  geom_bar(position = "dodge")
```

Remember, the two most commonly used Gestalt principles? Similarity &
proximity. This plot takes care of all that for us.

Position `"fill"` is also really nice but this provides the proportional
values, noticed that the Y axis name doesn't change, it still says
"count", So you'll need to use a `scale_*_*()` function, see below, to
change it.

```{r}
# position "fill"
g +
  geom_bar(position = "fill")
```

Did you see how it was convenient to use the object `g` Instead of
copying and pasting the base layer? We'll see this couple more times in
the tutorial.

## Adjusting aesthetics using the scales functions

Before we move onto the other layers the last key thing you should know
about the first three essential layers is how to adjust the scales. This
is done using one of the scales functions. Like the geom functions,
there is an intuitive naming convention. All the scale functions follow
the form `scale_*_*()`.

The first `*` refers to which scale you want to modify, e.g.:

-   `scale_x_*()`

-   `scale_y_*()`

-   `scale_color_*()`

    -   Also `scale_colour_*()`

-   `scale_fill_*()`

-   `scale_shape_*()`

-   `scale_linetype_*()`

-   `scale_size_*()`

This **must** match the aesthetics that you're actually using in the
plot, or nothing will happen.

The second `*` refers to which *type* of scale to use, e.g.:

-   **`scale_x_continuous()`**

-   **`scale_y_continuous()`**

-   **`scale_color_discrete()`**

    -   also, `scale_colour_discrete()`

In our scatterplot we have three aesthetics, X, Y and color. Let's take
at how to use these functions one at a time. Here's our starting point:

```{r}
# Base layers
p <- ggplot(mtcars, aes(wt, mpg, col = factor(cyl))) +
  geom_point(shape = 16, alpha = 0.6, size = 3)

p
```

Update colour scale, manual:

```{r}
# Scales (i.e. aesthetics)
p <- p +
  scale_colour_manual("Cylinders", values = myBlues)

p

```

Noticed that the first argument is the name that we want applied to that
scale. In this case we also use the values function to add the colour
palette that we defined in the previous section.

Update X axis, continuous:

```{r}
p <- p +
  scale_x_continuous("Weight (1000 lbs)", limits = c(0,6), breaks = 0:6, expand = c(0,0))
p

```

The `limits` argument defines the scale's range, `breaks` defines where
to place the major tick marks, and `expand` is a two-element long
numeric vector specifying the *additive* and *multiplicative* padding at
the end of the axes. The default setting here is 5% extra room of the
entire scale. The reason for this is that if you have a data point
directly on the scale, and you don't have the buffering, then it could
be obscured by the axis. If you're sure that you don't need buffering,
it's fine to remove it.

Update Y axis, continuous:

```{r}
# Scales (i.e. aesthetics)
p <- p +
  scale_y_continuous("mpg", limits = c(10,35), expand = c(0,0))
p
```

Of course you can do this all in one step, I broke it apart above so you
can see each piece.

```{r}
ggplot(mtcars, aes(wt, mpg, col = factor(cyl))) +
  geom_point(shape = 16, alpha = 0.6, size = 3) +
  scale_colour_manual("Cylinders", values = myBlues) +
  scale_x_continuous("Weight (1000 lbs)", limits = c(0,6), breaks = 0:6, expand = c(0,0)) +
  scale_y_continuous("mpg", limits = c(10,35), expand = c(0,0))
```

Notice that the way our plotting commands are built means we can comment
out a specific line if we don't need it.

```{r}
ggplot(mtcars, aes(wt, mpg, col = factor(cyl))) +
  geom_point(shape = 16, alpha = 0.6, size = 3) +
  scale_colour_manual("Cylinders", values = myBlues) +
  # scale_x_continuous("Weight (1000 lbs)", limits = c(0,6), breaks = 0:6, expand = c(0,0)) +
  scale_y_continuous("mpg", limits = c(10,35), expand = c(0,0))
```

This won't work if it's the last line that we want to comment out. We'd
have a hanging `+`, which would indicate to R that our command is not
complete. So, if you feel that you want to experiment and comment out
certain commands, you can end your ggplot call with `NULL`.

```{r}
ggplot(mtcars, aes(wt, mpg, col = factor(cyl))) +
  geom_point(shape = 16, alpha = 0.6, size = 3) +
  scale_colour_manual("Cylinders", values = myBlues) +
  scale_x_continuous("Weight (1000 lbs)", limits = c(0,6), breaks = 0:6, expand = c(0,0)) +
  # scale_y_continuous("mpg", limits = c(10,35), expand = c(0,0)) +
  NULL
```

You may be tempted to copy and paste code layers between plots, but be
very cautious with copying and pasting scale functions! If you apply a
limit to a data set, you may remove data that's actually interesting
just look at. Always plot the three essential layers *without* modifying
the scales, then apply these functions to clean up your plot.

# ggplot2: The optional 4 layers

Let's take a look at the last four layers.

## Layer 4: Statistics

We'll look at applying two very common statistics to your plots in this
layer.

### Smoothing lines

We can add a LOESS curve, which is a non-parametric model, using:

```{r}
p + 
  stat_smooth(se = FALSE) # LOESS, default
```

The `se` argument specifies whether we want to plot the *95% Confidence
Interval* or not.

We can plot a classic OLS model by specifying the `method`:

```{r}
# Add an OLS linear model
p + stat_smooth(method = "lm", se = FALSE) # Linear Model
```

Notice that because we have a `color` aesthetic, models are applied to
each group separately. The reason for this is that there is an invisible
aesthetic, called `group`, which inherits group identity from `color`.
We can override the `group` aesthetic by defining it as a *dummy*
variable, e.g. `1`, within the stats layer. Thus, we can color the
points according to one variable, but apply the model to *all* the data.

```{r}
# override the col/group aes:
p + stat_smooth(aes(group = 1), method = "lm", se = FALSE) # one group
```

### Summary statistics:

Other common statistics that you'll often apply summary statistics.
Before we proceed, let's clean up our data. I will use some base package
functions to convert `cyl` and `am` to proper factor variables.

```{r}
mtcars$fcyl <- as.factor(mtcars$cyl)
mtcars$fam <- as.factor(mtcars$am)
```

We saw that positions are defined with the `position` argument within an
geom, but we can also specify them using functions. As you would expect
these functions take the form `position_*()`. The two advantages of
defining positions in a function are that we can specify the arguments
precisely, and we can reuse positions in many layers or in many plots.

```{r}
posn_d <- position_dodge(0.3)
posn_j <- position_jitter(0.3)
posn_jd <- position_jitterdodge(0.3, dodge.width = 0.3)
```

All right let's begin to make a plot. Here's the base layer with 3
variables:

```{r}
q <- ggplot(mtcars, aes(fcyl, wt, col = fam))
q
```

First thing we want is to plot individual points.

```{r}
q +
  geom_point(alpha =  0.6)
```

You can see the case for jittering here. We can use `geom_jitter()` and
specify the width.

```{r}
q +
  geom_jitter(width = 0.3, alpha =  0.6)
```

But we can also use `geom_point()` and specify jittering using our
position function.

```{r}
q +
  geom_point(position = posn_j, alpha =  0.6)
```

Indeed, we may want to have not just jittering, but both jitter and
dodge.

```{r}
q +
  geom_point(position = posn_jd, alpha =  0.6)
```

Let's add some statistics: the mean and standard deviation. Before we
plot the mean and standard deviation, let's take a look at how we are
going to get them. I'll begin with some random numbers as an example:

```{r}
set.seed(136)
xx <- rnorm(100, 10, 6)
mean(xx)
sd(xx)
```

The function were going to use to calculate the mean and standard
deviation, `smean.sdl()`, comes from the `Hmisc` package. The `mult`
argument says we only want one multiple of the standard deviation. We
loaded this package at the beginning of the script.

```{r}
smean.sdl(xx, mult = 1)
```

I loaded the package because I wanted to show you how it works, but if
you want to do this in your own functions you don't need to explicitly
called `Hmisc`. That's because the `ggplot2` function we're going to
use, `mean_sdl()`, will call the `Hmisc` function for us.

```{r}
mean_sdl(xx, mult = 1)
```

Can you spot the difference in the output format? The ggplot2 function
returns a dataframe. Each variable is named after an aesthetic we need
to plot: the mean and the mean +/- 1 standard deviation, thus `ymin`,
`y`, and `ymax`. `ymin` and `ymax` work just like X, Y, color, etc. that
we saw before. To use this function inside our plot, use:

```{r}
q +
  stat_summary(fun.data = mean_sdl, 
               fun.args = list(mult = 1),
               position = posn_d)
```

There are 3 things to note inside here.

First, we specified to function without brackets, `mean_sdl`, to the
`fun.data` argument. This argument expects the output that we saw above
including all three aesthetics. If we only want to specify one, we can
use `fun.ymin`, `fun.y`, `fun.ymax` or any combination. You probably
guessed that the `fun` part means *function*.

Second, notice that `mult = 1` is necessary, and in `stat_summary()`, We
need to provide arguments to our stat function as a list to the
`fun.args` argument.

Third, we again set the position using a previously defined object.

Let's move onto inferential statistics: the mean and the 95% CI. The
function to use here is `mean_cl_normal()`. It works exactly like what
we saw earlier with `mean_sdl()`, except we don't need to specify any
arguments since it defaults to 95%. Although it has normal in the name,
it is the t-corrected 95% CI. In this case we only need to dodge.

```{r}
q +
  stat_summary(fun.data = mean_cl_normal,
               position = posn_d)
```

We can of course combine individual points and a summary:

```{r}
q +
  geom_point(position = posn_jd, alpha =  0.6) +
  stat_summary(fun.data = mean_sdl, 
               fun.args = list(mult = 1),
               position = posn_d)
```

## Layer 5: Coordinates

In the video course, I presented the `coord_` layer in the context of
the aspect ratio. Here I want to take a look at a very important
practical aspect. Continue with the plot we just made above, where we
had the mean and standard deviation:

```{r}
q +
  stat_summary(fun.data = mean_sdl, 
               fun.args = list(mult = 1),
               position = posn_d)
```

Imagine that I now want to set a limit on the Y axis, since the error
bars don't go above 5.

```{r}
q +
  stat_summary(fun.data = mean_sdl, 
               fun.args = list(mult = 1),
               position = posn_d) +
  scale_y_continuous(limits = c(1,5),
                     expand = c(0,0))
```

Compare the two plots above. Do you notice that the error bars are
different, and that we got a warning message? This is a serious concern!
Why did the error bars change? This is why I mentioned earlier that you
should never just apply scale functions without first seeing all of your
data.

The scale functions will *filter* data that is *outside* of the limits.
That means that all of the automatically calculated statistics, like the
mean and standard deviation, will be calculated *only* on that filtered
dataset. This may be what you want, but it's probably not.

What we really want to do is zoom in on the plot, but not filter the
data. With this we can use a coordinate function. As you probably
already expected these all have the form `coord_*()`.

```{r}
q +
  stat_summary(fun.data = mean_sdl, 
               fun.args = list(mult = 1),
               position = posn_d) +
  scale_y_continuous(expand = c(0,0)) +
  coord_cartesian(ylim = c(1,5))

```

## Layer 6 Facets, Small Multiples

Small multiples are implemented with the facet later. Again, they
following a nice naming convention. Each function follows the form
`facet_*()`. There are a couple ways in which we can input variables. I
tend to prefer formula notation: `y ~ x`, Which in this case becomes
`rows ~ columns`.

When you have a small number of groups in a categorical variable, you
can use `facet_grid()`. Let's begin with an example:

```{r}
m <- ggplot(mtcars, aes(wt, mpg)) +
  geom_point()

m
```

Split by rows:

```{r}
m + facet_grid(cyl ~ .)
```

Split by columns:

```{r}
m + facet_grid(. ~ am)
```

Split by both:

```{r}
m + facet_grid(cyl ~ am)
```

Noticed that cyl and am are numeric. `ggplot2` will convert these to
factors. If we want to set the order, we should make them factors
ourselves and properly set the order of the levels. See the Factors
chapter in the *Data Analysis with R* course for more details.

For many groups, use `facet_wrap()`:

```{r}
m + facet_wrap(~ gear, ncol = 2)
```

## Layer 7: Themes, Non-data Ink

The themes layer is where we modify all of the non-data ink on our plot.
Just like all the other layers the functions here take the form
`theme_*()`. For example, there are many built-in themes. Probably the
the most useful is:

```{r}
p +
  theme_classic()
```

But `theme()` allows us to modify any non-data ink element on the plot.
There are three types of non-data ink: text, lines and rectangles. Each
element gets an intuitive name, using a `.` notation. The name tells us
which level of the hierarchy of specific element falls into. All of
these names are arguments in `theme()`. For example, for text elements:

    theme(
    text, 
      axis.title, 
        axis.title.x,
          axis.title.x.top, 
          axis.title.x.bottom, 
        axis.title.y, 
          axis.title.y.left,
          axis.title.y.right, 
      title, 
        legend.title, 
        plot.title,
        plot.subtitle, 
        plot.caption, 
        plot.tag, 
      axis.text, 
        axis.text.x, 
          axis.text.x.top,
          axis.text.x.bottom, 
        axis.text.y, 
          axis.text.y.left, 
          axis.text.y.right,
      legend.text,
      strip.text, 
        strip.text.x, 
        strip.text.y)

For line elements:

    theme(
    line,
      axis.ticks,
        axis.ticks.x, 
          axis.ticks.x.top,
          axis.ticks.x.bottom,
        axis.ticks.y,
          axis.ticks.y.left, 
          axis.ticks.y.right,
        axis.line,
          axis.line.x, 
            axis.line.x.top,
            axis.line.x.bottom, 
          axis.line.y,
            axis.line.y.left, 
            axis.line.y.right, 
      panel.grid,
        panel.grid.major, 
          panel.grid.major.x, 
          panel.grid.major.y, 
        panel.grid.minor,
          panel.grid.minor.x,
          panel.grid.minor.y)

And for rectangle elements:

    theme(
    rect,
      legend.background, 
      legend.key,
      legend.box.background, 
      panel.background,
      panel.border,
      plot.background, 
      strip.background, 
        strip.background.x, 
        strip.background.y)

To modify an element, just call its argument in `theme()` and apply the
modification using either `element_text()`, `element_ _line()`, or
`element_rect()`, depending on the element you are modifying. e.g.

```{r}
p +
  theme(axis.text = element_text(colour = "red"))
```

There is one other element function that we haven't discussed yet:
`element_blank()`. We can use this in a plot to remove any item. That
is, it won't be drawn at all. In this example I set all rectangles to
blank:

```{r}
p +
  theme_classic() +
  theme(rect = element_blank())
```

Actually, in addition to non-data ink we also modify the position of
specific elements on the plot, like margins and spacing, in the themes
layer. The most common modification you may make is to change the
position of the legend using the `legend.position` argument. This takes
a character argument, e.g. `"none"` to remove the legend, or `"top"`,
`"right"`, `"bottom"`, `"left"` to center it on that side of the plot.

```{r}
p +
  theme(legend.position = "bottom")
```

You can position the legend inside the plot by specifying a two-element
long numeric vector. Specifies the X and Y position of the legend in
`npc` (Normalised Parent Coordinates) units. These values range from 0
to 1, where 0 is the min and 1 is the max value of the axis.

```{r}
p +
  theme_classic() +
  theme(legend.position = c(0.85, 0.75))
```

Of course we can combine this altogether:

```{r}
p +
  theme_classic() +
  theme(rect = element_blank(),
        axis.text = element_text(colour = "black"),
        legend.position = c(0.85, 0.75))
```

# Saving plots

Broadly speaking, there are two types of images: raster & vector.

Raster images have pixels & resolution. Think of the images your camera
takes. Typical formats include: - jpg - png - tif - gif - bmp

Vector images don't actually contain an image, they just have the
instructions to make images. Typical formats include: - pdf - ai - svg -
ps - eps

To save the last active image just run `ggsave()` after you print it to
the screen.

```{r eval = FALSE}
ggsave("myPlot.png", width = 6, height = 6, units = "in")
ggsave("myPlot.pdf", width = 6, height = 6, units = "in")
```

The extension on the file name will tell `ggsave()` what format to use
to save the image. If we don't specify a path, it will be saved to the
working directory. That's why it's always good to work inside of a
project in RStudio.

If you want to explicitly say which plot to save just specify it as the
second argument.

```{r eval = FALSE}
ggsave("myPlot_p.png", p, width = 6, height = 6, units = "in")
ggsave("myPlot_p.pdf", p, width = 6, height = 6, units = "in")
```
