---
title: "Diamonds exercises"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

## The `diamonds` Dataset

Exercise 9.1 (Examine structure)

- What type of data is contained in each column?

- Use some of the functions we introduced in the dataframes section to explore the basic structure of our new object.

The dataset contains the following variables:

```{r ex9_1, echo=FALSE}
# Variable names
# names(diamonds)

# A compact view
glimpse(diamonds)

# Summary statistics
# summary(diamonds)
```

Our data:

```{r}
diamonds
```

## Exercise 9.2 (Counting individual groups)

How many diamonds with a clarity of category "IF" are present in the dataset?

```{r}
(num_IF <- sum(diamonds$clarity == "IF"))

diamonds %>% 
  group_by(clarity) %>% 
  count()

table(diamonds$clarity)

table(diamonds$clarity, diamonds$color)
```

A: From a total of `r nrow(diamonds)` diamonds, there are `r num_IF` IF clarity diamonds in our dataset.

## Exercise 9.3 (Measuring proportions)

- What proportion (0-1) (or percentage (0-100)) of the total do they represent?

```{r}
num_IF / nrow(diamonds) # proportion

num_IF / nrow(diamonds) * 100 # percent
```

## Exercise 9.4 (Summarizing proportions)

- What proportion of the whole is made up of each category of clarity?

```{r echo=FALSE}

diamonds %>%
  count(clarity) %>%
  mutate(
    # n_ = n(), # 8, n from input, not diamons
    # n_row = nrow(diamonds),
    prop = n/nrow(diamonds)) %>% 
  knitr::kable(caption = "The proportion of each type of clarity diamond.")

# Longer form:
diamonds %>% 
  group_by(clarity) %>% 
  summarise(n = n()/nrow(diamonds)) 
  
# The intersection of two variables (as a data.frame)
# with grouping by clarity
diamonds %>% 
  group_by(clarity, color) %>% 
  summarise(n = n())

# A contigency table (not a data.frame object)
table(diamonds$clarity, diamonds$color)

table(diamonds$clarity) / length(diamonds$clarity)
table(diamonds$clarity) / nrow(diamonds)

```

## Exercise 9.5 (Finding prices I)

- What is the cheapest diamond price overall?

```{r}
min(diamonds$price)
```

- What are the properties of the cheapest and the most expensive diamonds?

```{r}
which.min(diamonds$price)

# use to find the rest of the values:
# Hard coding :(
diamonds %>% 
  slice(1)

# Integrated
# By index position
diamonds %>% 
  slice(which.min(price), which.max(price)) %>%
  # select(-price) %>%
  print()

# By logical vector
diamonds %>%
  filter(price == min(price) | price == max(price)) %>%
  # select(-price) %>% 
  print()

# Although it worked, try to not use !price to _not_ select
# a column. Use it for negation in logical expressions:
diamonds %>% 
  filter(price < 500)
# The negation
diamonds %>% 
  filter(!(price < 500))
```

## Exercise 9.6 (Finding prices I)

- What is the range (single number, Q4-Q0) of diamond prices?

```{r}
# As a vector
diff(range(diamonds$price))

# As a data.frame (tibble)
diamonds %>%
  summarise(singleRange = max(price) - min(price))
```

- The min & max, separately, for each category of cut and color? How would you describe what the results show? (kind of difficult)

```{r}
table(diamonds$cut, diamonds$color)

# 4 variables: cut, color, min, max
# dim: 35 x 4
diamonds %>% 
  group_by(cut, color) %>% 
  summarise(min = min(price),
            max = max(price)) -> dia_range_2_col
dia_range_2_col

# pivot, specity var to exclude
dia_range_2_col %>% 
  pivot_longer(-c(cut, color))

# pivot, specity var to include
dia_range_2_col %>% 
  pivot_longer(c(min, max))

# 3 variables cut, color, range
# dim: 70 x 3
diamonds %>% 
  group_by(cut, color) %>% 
  summarise(range = range(price)) -> dia_range_1_col
dia_range_1_col

# Add a column for defining which value is max and min
dia_range_1_col %>% 
  mutate(metric = c("min", "max")) -> dia_range_1_col_ID
```

- Produce a plot from the previous exercise. Does this change how you would summarize the previous result?

y ~ x plots are typically described as dependent variable ~ independent variable, $y = f(x)$ 

Ideas:
- A bar plot: where for entry, we have two bars one min and one vor max
  - x-axis: categorical, here cut or color, or both
  - y-axis: continuous, here min or max, or both
- A box-plot: i.e. using the entire dataset. (will see later)
- A heat map (categorical x & y): x = cuts, y = colors, intersection is colored according to the value of a continuous variable (i.e. min or max). (e.g. )

```{r}
# Define position for dots in plots:
posn_d <- position_dodge(width = 0.25)
# But can also use position = "dodge"

# Define the data and aesthetics  
ggplot(dia_range_1_col, aes(cut, range, color = color)) + geom_point(position = posn_d) +
  # geom_path() +
  # geom_line() +
  # scale_color_brewer() +
  NULL

# try with 2 col, but we'd need two ploting functions:
ggplot(dia_range_2_col, aes(cut, min, color = color)) +
  geom_point(position = posn_d) +
  NULL
ggplot(dia_range_2_col, aes(cut, max, color = color)) +
  geom_point(position = posn_d) +
  NULL

# Use IDs for each value:
ggplot(dia_range_1_col_ID, aes(cut, range, color = color)) +
  geom_point(position = posn_d) +
  # geom_path() +
  # geom_line() +
  facet_grid(metric ~ ., scales = "free_y") +
  NULL


# Avoid bar plots in this case
# More appropriate for absolute counts
# geom_jitter() # add random noise on x
# geom_col(color = "red") # default position = "stack"
# We can define aes in a geom_ but it doesn't help here:
  # geom_point(aes(y = max)) +
```

## Exercise 9.7 (Finding prices II)

- What is the average diamond price in each category of cut and color?

```{r}

```

- Plot the raw values for each of the above groups _and_ the mean and standard deviation.

```{r}

```

- Are boxplots useful when representing this data?

```{r}

```

## Exercise 9.8 (Plotting)

- Produce a plot that shows the relationship between the size of a diamond (as measured by carat) and the price. 

```{r}

```

- Is a transformation of either axis in order? How would you apply it?

```{r}

```

- How can we distinguish between different groups of categorical variables, such as cut, clarity or color? 

```{r}

```

- How about other continuous variables? 

```{r}
diamonds %>% 
  group_by(cut, color) %>% 
  summarise(min = min(price),
            max = max(price)) %>% 
  # data.frame() %>% 
  
  print()
```
