---
title: "Martian-exercises"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)


# Import data for interactive mode
# martians <- read_tsv("00-data/martians.txt")

# Import data when document is rendered
martians <- read_tsv("../00-data/martians.txt")
```

Above, we compensated for different environments and working directories in interactive mode and during document rendering. **This is not a good solution** in the long terem... see the `{here}` package.

## Exercise 9.8 (Only tall martians)

Use the entire dataset for this exercise, i.e. Site I and Site II. Let the median divide the height variable into lower and upper halfs. Calculate the proportion of blue and red-nosed Martians in each half. Do you think there is a real difference in the proportion of nose colors between the lower and upper 50% of the sample (according to height)?


```{r}
martians %>% 
  select(Height, Nose) %>% 
  mutate(Short = Height < median(Height)) %>%  # define the two parts of the dataset
  group_by(Short, Nose) %>% 
  summarise(count_nose = n())  %>%
  # ungroup() %>%
  # group_by(Short) %>%
  mutate(prop = prop.table(count_nose)) %>% 
  ggplot(aes(Short, prop, fill = Nose)) +
  geom_col() +
  scale_fill_manual(values = c("Blue" = "light blue", "Red" = "pink"))

# Convenient way to see the groups in a grouped dataframe as a list of dataframes
# group_split()
```

Can I get `ggplot2` to calculate the absolute numbers and proportions of groups?

We can use `geom_bar()` to calculate from the raw values, instead of`geom_col()`, which uses pre-calculated values. 

```{r}
# y = Height
# x = Site
H_med <- median(martians$Height)

table(martians$Height < H_med, martians$Nose)

ggplot(martians, aes(Height < H_med, fill = Nose)) +
  geom_bar() +
  scale_fill_manual(values = c("Blue" = "light blue", "Red" = "pink")) +
  NULL

```




## 